---
title: "EMA, Grazia Mirabito"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r, cleaning data}
knitr::opts_chunk$set(echo = TRUE)
#import packages
library (ggplot2)
library(dplyr)
library(psych)
library(plyr)
library(doBy)
library(lme4)
library(jtools)

#import data
data <- readxl::read_xlsx("~/Desktop/Thesis/Aggregating-EMA/Long-Format.xlsx")

#recodedata
data$PhysicalActivity<- recode(data$PhysicalActivity, "YES"= 1, "NO"= 2)

#look at stats before aggregating
#describe(data)

#parsing the timestamp string to just have the date mm/dd
data$DATE <- substr(data$TIMESTAMP, 1, 10)

#sort new data and aggregate #getting average of scores by date and id 
new.data1 <- summaryBy(GROUP + TMS + Curiousity + Decentering + STRESSOR + PROMIS + PWB + Rumination + SLEEP + PhysicalActivity + FORMAL + INFORMAL ~ DATE + ID, data=data, FUN=mean, na.rm=T)

#make date numeric and not a string
new.data1$DATE <- as.Date(new.data1$DATE, "%m/%d/%Y")

describe(new.data1)
#exporting file
library("writexl")
#write_xlsx(new.data1,"~/Desktop/average-Long-Format.xlsx")
```

``` {r, cronbach data cleaning}
chron <- readxl::read_xlsx("~/Desktop/Thesis/Aggregating-EMA/TOGETHER-EMA.xlsx")

#getting the last data point for the morning survey 
chron.new <- unique(chron$ID)
a<-length(chron.new) 
last <- c()
for (i in 1:a) {
  temp<-subset(chron, ID==chron.new[i])
  if (dim(temp)[1] > 1) {
    last.temp<-temp[dim(temp)[1],]
  }
  else {
    last.temp<-temp
  }
  last<-rbind(last, last.temp)
}

as.data.frame(last)

#getting the first data point for the morning survey 
t.first <- chron[match(unique(chron$ID), chron$ID),]
as.data.frame(t.first)

#exporting 
#write_xlsx(last,"~/Desktop/average-last.xlsx")
#write_xlsx(t.first,"~/Desktop/average-first.xlsx")

```


```{r, empty model}
#cleaned data in excel to get day of study numbers 
#importing new data 
EMA <- readxl::read_xlsx("~/Desktop/Thesis/Aggregating-EMA/EMA-DATA-PREFINAL.xlsx")
#recode group to 0 and 1 
#0= control 1= intervention
EMA$GROUP.mean <- recode(EMA$GROUP.mean, "1"=0 , "2"=1 )
#reode day to start with 0-34
EMA$TIME <- (EMA$DAY-1)
EMA$ID <- substr(EMA$ID, 13, 15)
as.numeric(EMA$ID)
EMA$PROMIS<- round(EMA$PROMIS.mean)
EMA$STRESSOR<- round(EMA$STRESSOR.mean)
EMA$PWB <- round(EMA$PWB.mean)
EMA$Rumination <- round (EMA$Rumination.mean)
EMA$GROUP <- round(EMA$GROUP.mean)
#exporting 
#write_xlsx(EMA,"~/Desktop/EMA-final.xlsx")
#subset by averaging data per day for intervention 
#daybyday.data <- summaryBy(TMS.mean + Curiousity.mean + Decentering.mean + STRESSOR + PROMIS + PWB + Rumination + SLEEP.mean + PA + FORMAL.mean + INFORMAL.mean ~ TIME2 + GROUP, data=EMA.subset, FUN=mean, na.rm=T)

new.daybyday <- subset(daybyday.data, GROUP <= 1)
```

```{r}
#compliance 
compliance_stats <-  ddply(EMA.subset, "ID", summarize,
                           num_days = length(unique(TIME2)),
                           num_obs = length(ID))

describe(compliance_stats)
table(compliance_stats$num_days, useNA = "no")
table(compliance_stats$num_obs, useNA = "no")
#histogram
ggplot(data = compliance_stats, aes(x = num_days)) +
  geom_histogram(fill="white", color="black") + 
  labs(x = "Number of Observations Completed")
#280 data points missing out of 3135 
```


```{r}
#all from session c in longitudinal folder. 
#visualize for groups
ggplot(data = EMA[which(EMA$ID < 20),],
       aes(x = TIME, group=ID), color=factor(ID)) +
  guides(color="none") + #to suppress guide
  # first variable
  geom_line(aes(y=TMS.mean, color=factor(ID)))

#plot more variables for one person 
ggplot(data = EMA[which(EMA$ID == 105), ],
       aes(x = TIME, group= ID)) +
  # first variable
  geom_line(aes(y=TMS.mean, colour="red")) +
  geom_point(aes(y=TMS.mean, colour="red")) +
  # second variable
  geom_line(aes(y=Rumination.mean, color="blue")) +
  geom_point(aes(y=Rumination.mean, color="blue")) 

#plot a few people
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=TMS.mean),colour="#F27781") +
  geom_point(aes(y=TMS.mean),colour="#F27781") +
  facet_wrap(~ID, nrow=5)

#plot a few people with 2 variables 
cols <- c("TMS"="#F27781", "Rumination"="#18298C", "Stress"= "blue","PWB"= "green", "Depression"= "yellow", "Sleep"= "black", "Physical Activity"= "brown" )
ggplot(data = EMA.subset[which(EMA$ID <= 104),  ],
       aes(x = TIME2, group= factor(ID))) +
  geom_line(aes(y=TMS.mean, colour="TMS")) +
  geom_point(aes(y=TMS.mean, colour="TMS")) +
  geom_line(aes(y=STRESSOR, color="Stress")) +
  geom_point(aes(y=STRESSOR, color="Stress")) +
  geom_line(aes(y=Rumination, color="Rumination")) +
  geom_point(aes(y=Rumination, color="Rumination")) +
  geom_line(aes(y=PWB, color="PWB")) +
  geom_point(aes(y=PWB, color="PWB")) +
  geom_line(aes(y=PROMIS, color="Depression")) +
  geom_point(aes(y=PROMIS, color="Depression")) +
  geom_line(aes(y=SLEEP.mean, color="Sleep")) +
  geom_point(aes(y=SLEEP.mean, color= "Sleep")) +
  geom_line(aes(y=PhysicalActivity.mean, color="Physical Activity")) +
  geom_point(aes(y=PhysicalActivity.mean, color= "Physical Activity")) +
  facet_wrap(~ID, nrow=5)
```


```{r}
#mindfulness
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=TMS.mean),colour="#F27781") +
  geom_point(aes(y=TMS.mean),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Mindfulness") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Mindfulness over Time")
#rumination 
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=Rumination),colour="#F27781") +
  geom_point(aes(y=Rumination),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Rumination") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Rumination over Time")
#sleep 
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=SLEEP.mean),colour="#F27781") +
  geom_point(aes(y=SLEEP.mean),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Sleep") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Sleep over Time")
#PA 
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=PhysicalActivity.mean),colour="#F27781") +
  geom_point(aes(y=PhysicalActivity.mean),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Physical Activity") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Physical Activity over Time")
#Depression 
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=PROMIS),colour="#F27781") +
  geom_point(aes(y=PROMIS),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Depression") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Depression over Time")
#PWB
ggplot(data = EMA[which(EMA$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=PWB),colour="#F27781") +
  geom_point(aes(y=PWB),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("PWB") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in PWB over Time")
#stress
ggplot(data = EMA.subset[which(EMA.subset$ID <= 106), ],
       aes(x = TIME, group= factor(ID))) +
  geom_line(aes(y=STRESSOR),colour="#F27781") +
  geom_point(aes(y=STRESSOR),colour="#F27781") +
  facet_wrap(~ID, nrow=5)+
  xlab("Time") + ylab("Stress") +
  theme_classic() +
  theme(axis.title=element_text(size=12),
        axis.text=element_text(size=12),
        plot.title=element_text(size=12, hjust=.5)) +
  ggtitle("Change in Stress over Time")

```


```{r}
EMA.tx<- subset.data.frame(EMA, GROUP==1)
EMA.control<- subset.data.frame(EMA, GROUP==0)
```

```{r}
#mindfulness change over time between intervention and control
#running uncentered TMS
mind.model1 <- lmer(TMS.mean ~ TIME*GROUP + (1 + TIME | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(mind.model1)
summ(mind.model1)
#uncentered curiousity
mind.model2<- lmer(Curiousity.mean ~ TIME*GROUP + (1 + TIME | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(mind.model2)
#uncentered decentering
mind.model3 <- lmer(Decentering.mean ~ TIME*GROUP + (1 + TIME | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(mind.model3)

#plot TMS Subscales
ggplot(data=EMA, aes(x=TIME, y=Curiousity.mean, group=factor(GROUP), colour="gray"), legend=TRUE) +
  geom_smooth(aes(group=ID), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=.5, color="gray") +
  geom_smooth(data=EMA.control, aes(x=TIME, y=Curiousity.mean), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=2, color="blue", show.legend= TRUE) +
  geom_smooth(data=EMA.tx, aes(x=TIME, y=Curiousity.mean), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=2, color="red", show.legend= TRUE) +
  xlab("TIME") + ylab("Mindfulness") +
  theme_classic() +
  theme(axis.title=element_text(size=18),
        axis.text=element_text(size=14),
        plot.title=element_text(size=18, hjust=.5)) +
  ggtitle("Within-Person Association\nCuriousity and TIME")


ggplot(data=EMA, aes(x=TIME, y=Decentering.mean, group=factor(GROUP), colour="gray"), legend=TRUE) +
  geom_smooth(aes(group=ID), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=.5, color="gray") +
  geom_smooth(data=EMA.control, aes(x=TIME, y=Decentering.mean), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=2, color="blue") +
  geom_smooth(data=EMA.tx, aes(x=TIME, y=Decentering.mean), method=lm, se=FALSE, fullrange=FALSE, lty=1, size=2, color="red") +
  xlab("TIME") + ylab("Mindfulness") +
  theme_classic() +
  theme(axis.title=element_text(size=18),
        axis.text=element_text(size=14),
        plot.title=element_text(size=18, hjust=.5)) +
  ggtitle("Within-Person Association\nDecentering and TIME")

#only in tx group
mind.tx <- lmer(Decentering.mean ~ TIME + (1 + TIME | ID), data=EMA.tx, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(mind.tx)
```
**This shows us that for the intercept the average TMS score was 21.24 for the control group... and the B value for group.mean shows that the intial level (average) for the interention is 21.24 + 0.67 at about 21.93 on day 0. B value DAY2 at 0 shows the control group change in minduflness over time. and the interaction shows how much higher the rate of change is for the treatment group compared to the control group at 0.13... so the intervention group had a 0.13 change in mindfulness over the 5 weeks compared to control group which had 0.** 

**The random component shows that for the random intercept has a SD=4.78 which indivats that 95% of the population vary between +/- 9.56 units of the typical intercept for their group. The slope random component has a SD= 0.14 which shows that 95% of the poplulation nvary between +/- 0.28 units of the typical slope for it's group.**

**The intercepts and slopes covary at -0.19, this means that people that have a lower intercepts so lower TMS subscale scores have a more strongly positive effect of TMS by day. Therefore, the relationship of time on TMS score is more steep or even becomes more strongly positive as you decrease Intial TMS scores. **

```{r}
#I think just repeat for all the other outcome var (rumination, sleep, exercise, anxiety, stress, well-being, mood)
#rumination
rumination.model1 <- lmer(Rumination.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(rumination.model1)
#sleep 
sleep.model1 <- lmer(SLEEP.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(sleep.model1)
#exercise 
exercise.model1 <- lmer(PhysicalActivity.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(exercise.model1)
#anxiety and stress
anxiety.model1 <- lmer(STRESSOR.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(anxiety.model1)
#well-being
PWB.model1 <- lmer(PWB.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(PWB.model1)
#mood 
mood.model1 <- lmer(PROMIS.mean ~ TIME*GROUP + (1 + DAY2 | ID), data=EMA, REML= TRUE, control=lmerControl(optimizer="bobyqa"))
summary(mood.model1)
```

```{r} 
#ICC 
#level 1== day 
#level 2== people 
#level 3== group (control vs intervention)

#get ICC of within a person dependence for mindfulness
Mindfulness<- lmer(TMS.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Mindfulness)

#ICC of within group dependence for curiousity
Curiousity <- lmer(Curiousity.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Curiousity)

#ICC of within group dependence for decentering
Decentering <- lmer(Decentering.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Decentering)

#ICC of within group dependence for stressor
Stressor <- lmer(STRESSOR.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Stressor)

#ICC of within group dependence for promis
Promis <- lmer(PROMIS.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Promis)

#ICC of within group dependence for well being
PWB <- lmer(PWB.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(PWB)

#ICC of within group dependence for rumination
Rumination <- lmer(Rumination.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Rumination)

#ICC of within group dependence for sleep
Sleep <- lmer(SLEEP.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(Sleep)

#ICC of within group dependence for PA
PA <- lmer(PhysicalActivity.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
summ(PA)
```
**I ran multiple empty models of time on outcome variables by participant ID, the ICC's for within a person ranged from 0.25-0.74, making the DE range from 30-86.84, therefore, multilevel modelling would be good to do to account for the nesting of the data.** 


```{r, run stats INCOMPLETE} 
#running stats 
#center variables 
library(robumeta)
#EMA$TMS.center <- group.center(EMA$TMS.mean, EMA$ID) 
#run model of change of TMS over time.
#Ran.Slopes <- lmer(TMS.mean ~ 1 + (1 |ID), data = EMA,REML = TRUE)
#summary(Ran.Slopes)
#summ(Ran.Slopes)

```





