---
title: "Mediation-part1"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (ggplot2)
library(dplyr)
library(psych)
library(plyr)
library(doBy)
library(lme4)
library(nlme)
library(jtools)
library("broom.mixed")
library(misty)
#cleaned data in excel to get day of study numbers 
#importing new data 
EMA <- readxl::read_xlsx("~/Desktop/Thesis/Aggregating-EMA/EMA-DATA-PREFINAL.xlsx")
#recode group to 0 and 1 
#0= control 1= intervention
EMA$GROUP.mean <- recode(EMA$GROUP.mean, "1"=0 , "2"=1 )
#reode day to start with 0-34
EMA$TIME <- (EMA$DAY-1)
EMA$ID <- substr(EMA$ID, 13, 15)
as.numeric(EMA$ID)
EMA$PROMIS<- round(EMA$PROMIS.mean)
EMA$STRESSOR<- round(EMA$STRESSOR.mean)
EMA$PWB <- round(EMA$PWB.mean)
EMA$Rumination <- round (EMA$Rumination.mean)
EMA$GROUP <- round(EMA$GROUP.mean)
#subset
EMA.subset <- subset.data.frame(EMA, TIME>=7)
EMA.subset$TIME2 <- (EMA.subset$TIME-7)
EMA.tx<- subset.data.frame(EMA.subset, GROUP==1)
EMA.control<- subset.data.frame(EMA.subset, GROUP==0)
```

```{r}
#clean data and import packages
library(ggplot2); theme_set(theme_bw()) #for data viz
library(lme4) #for multilevel models
library(nlme) #for multilevel models
library(psych) #for describing data
library(tidyr); library(dplyr) #for manipulating data 
library(bmlm) #for Bayesian 1-1-1 mediation
library(reshape2)
library(broom) #for converting results summary to df
# Examine first few rows of the data set
head(EMA.subset, 30)

#part 1
#x= mindfulness 
#y= stress or well-being or mood 
#m= rumination 

#part 2 
#x= mindfulness 
#y= stress or well-being or mood 
#m= sleep

#part 3
#x= mindfulness 
#y= stress or well-being or mood 
#m= exercise


#need to center variables within a person
EMA.subset$TMS.C<- center(EMA.subset$TMS.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Curiousity.C<- center(EMA.subset$Curiousity.mean, type = "CWC", group = EMA.subset$ID)
#used to check mean of ID
#describeBy(EMA.subset$TIME2, group= EMA.subset$ID==10)
EMA.subset$Decentering.C<- center(EMA.subset$Decentering.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$PROMIS.C<- center(EMA.subset$PROMIS, type = "CWC", group = EMA.subset$ID)
EMA.subset$Stress.C<- center(EMA.subset$STRESSOR, type = "CWC", group = EMA.subset$ID)
EMA.subset$PWB.C<- center(EMA.subset$PWB, type = "CWC", group = EMA.subset$ID)
EMA.subset$Rumination.C<- center(EMA.subset$Rumination, type = "CWC", group = EMA.subset$ID)
EMA.subset$Sleep.C<- center(EMA.subset$SLEEP.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$PhysicalActivity.C<- center(EMA.subset$PhysicalActivity.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Formal.C<- center(EMA.subset$FORMAL.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Informal.C<- center(EMA.subset$INFORMAL.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Time.C<- center(EMA.subset$TIME2, type = "CWC", group = EMA.subset$ID)


#make x,y,m 
#EMA.subset$X.M <- EMA.subset$Curiousity.C
#EMA.subset$M.R <- EMA.subset$Rumination.C
#EMA.subset$Y.W <- EMA.subset$PWB.C

#variables of interest for part 1 
vars <- c("TMS.C", "Curiousity.C", "Decentering.C", "Stress.C", "PWB.C", "PROMIS.C", "Rumination.C")
#descriptives
describe(EMA.subset[,vars])

#PART 1 restructure for just mindfulness--> rumination --> stress
MRS.data <- EMA.subset[,c("GROUP", "ID","TIME2","Time.C","Curiousity.C","Stress.C", "Rumination.C")]
#look at  data set
#head(MRS.data)
#tail(MRS.data)

#ordering the data set to sort by ID and Time to make sure lead functions sort correctly by group and timing. 
MRS.data1<- MRS.data[order(MRS.data$ID, MRS.data$TIME2),]

#create lead (day1 curiosity, predictis day 2 rumiantion and day 3 well-being)
#creating lagged column for Ruminationi + 1 day and well-being + 2 days 
MRS.data.redo <- MRS.data1 %>%                            # Add lagged column
  group_by(ID) %>%
  dplyr::mutate(Rumination.lag = dplyr::lead(Rumination.C, n = 2, default = NA), Stress.lag = dplyr::lead(Stress.C, n = 2, default = NA)) %>% 
  as.data.frame()
MRS.data.redo 

#adding in the dummy coding variables with the lagged and centered variables required 

MRS.data.redo$X.M <- MRS.data.redo$Curiousity.C
MRS.data.redo$M.R <- MRS.data.redo$Rumination.lag
MRS.data.redo$Y.W <- MRS.data.redo$Stress.lag

#removing rumination and pwb to only leave the lagged variables in 
df = subset(MRS.data.redo, select = -c(Rumination.C, Stress.C) )


#melting the daya to use the dummy code double indicators
datalong.part1 <- reshape2:: melt(data=df,
                 id.vars=c("GROUP", "ID","TIME2","Time.C","Curiousity.C","Rumination.lag","Stress.lag","X.M"),
                 na.rm=FALSE, variable.name="dv",
                 value.name="z")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#adding the double indicators 
datalong.part1$dy <- ifelse(datalong.part1$dv=="Y.W", 1, 0)
datalong.part1$dm <- ifelse(datalong.part1$dv=="M.R", 1, 0)
datalong.part1$dvnum <- ifelse(datalong.part1$dv=="M.R", 1, 0)

#reordering for convenience 
datalong.part1 <- datalong.part1[order(datalong.part1$ID,datalong.part1$Time.C,datalong.part1$dm),]


```


```{r}
# NEED TO RUN THE MODEL_LME.PART1 NO MODERATION FIRST !!!!!!!

#meditation FIRST
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation model
model_lme.part1 <- lme(fixed = z ~ -1 + 
                             dm + dm:Curiousity.C + dm:Time.C + #m as outcome
                             dy + dy:Rumination.lag + dy:Curiousity.C + dy:Time.C, #y as outcome
                  random = ~ -1  +  dm:Curiousity.C + dy:Rumination.lag + dy:Curiousity.C | ID, correlation= corAR1(),
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.part1,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.part1)
```

Intercepts:
	0.0000 = intercept in the m as outcome model (d_Mi: We expect this to be zero because we person-centered all the data)

	0.0001 = intercept in the y as outcome model (d_Yi: We expect this to be zero (or extremely close) because we person-centered all the data)
Effects on M (Rumination):
	-0.095 = effect of x –> m (a: Mindfulness predicting Rumination) ***
	-0.007 = effect of time –> m (time trend in Rumination) ***
Effects on Y (Stress):
	-0.35 = effect of m –> y (b: Rumination predicting Well Being) ***
	0.124 = effect of x –> y (c’: Mindfulness predicting Stress, after adjusting for Rumination) ***
	0.0096 = effect of time –> y (time trend in Stress) ***

```{r}
#analyzing the mediation. 
#pulling out fixed effects info
FE.med <- fixef(model_lme.part1)
FE.med

#making parameter objects
a.med <- as.numeric(FE.med[3])
a.med
b.med <- as.numeric(FE.med[5])
b.med
cprime.med <- as.numeric(FE.med[6])
cprime.med

### Random Effects
#pulling out random effects info
VarCorr(model_lme.part1)

#The *variance* of the *a* paths
sig2_a <- as.numeric(VarCorr(model_lme.part1)["dm:Curiousity.C", "Variance"])
sig2_a

#variance of the *b* paths
sig2_b <- as.numeric(VarCorr(model_lme.part1)["dy:Rumination.lag", "Variance"])
sig2_b

#variance of the *c* paths
sig2_cprime <- as.numeric(VarCorr(model_lme.part1)["Curiousity.C:dy", "Variance"])
sig2_cprime

#The residual *variance* of the mediator variable, 

sig2_em <- (1.00*as.numeric(VarCorr(model_lme.part1)["Residual", "StdDev"]))^2
sig2_em

#The residual *variance* of the outcome variable, 
sig2_ey <- (0.6968058  *as.numeric(VarCorr(model_lme.part1)["Residual", "StdDev"]))^2
sig2_ey

#The *covariance* between the $a_{j}$ and $b_{j}$ paths 
#We need to convert the correlation to a covariance, which we can by using a `cortocov` function. We can just feed in the correlation coefficient and the two variances, and it will return a covariance. 

cortocov <- function (r, var1, var2) {
  cov=r*((var1*var2)^0.5)
  return(cov)
}

covajbj <- cortocov(r = as.numeric(VarCorr(model_lme.part1)["dy:Rumination.lag","Corr"]),
         var1 = as.numeric(VarCorr(model_lme.part1)["dm:Curiousity.C","Variance"]),
         var2 = as.numeric(VarCorr(model_lme.part1)["dy:Rumination.lag","Variance"]))

#indirect effect (mediated effect)
indirecteffect <- a.med*b.med + covajbj   
indirecteffect

#calculating c the total effect
totaleffect <- cprime.med + a.med*b.med + covajbj 
totaleffect

#percent mediated
percentmediated <- 100*(indirecteffect/totaleffect)  
percentmediated

#percent covariance
percentcovariance <- 100*(covajbj/indirecteffect) 
percentcovariance
```

```{r}
#starting moderation set up 
re <- data.frame(ranef(model_lme.part1))
re$ID <- row.names(re)
#get group in there
group <- as.data.frame(datalong.part1$GROUP,datalong.part1$ID)
names(group)[1] <- "GROUP"
group$ID <- row.names(group)
group.re <- summaryBy(GROUP ~ ID, data=group, FUN=mean, na.rm=T)
names(group.re)[2] <- "GROUP"
re$GROUP <- group.re$GROUP 
re$GROUP <- recode(re$GROUP, "0"=-0.5 , "1"=0.5 )

#don't think this line next is important? 
#re <- re %>% arrange(dm.TMS.C)

# add this new variable to the original data, resulting in a new dataframe, datalong2
datalong2.part1 <- merge(datalong.part1, dplyr::select(re, ID, GROUP), by = "ID")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#REDO for PART 1 restructure for mindfulness--> rumination --> well-being
#REDO for PART 1 restructure for mindfulness--> rumination --> mood
```

```{r}
#running moderation Model for Part 1 stress
#part 1
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)
#lme mediation and moderation
model_lme_mod.part2 <- lme(fixed = z ~ -1 + 
                       # m as outcome
                       dm + dm:Curiousity.C + dm:Time.C + 
                       # add in group as main effect and moderator
                       dm:GROUP.y + dm:Curiousity.C:GROUP.y + 
                       
                       # y as outcome
                       dy + dy:Rumination.lag + dy:Curiousity.C + dy:Time.C +
                       #  add in group as main effect and moderator
                       dy:GROUP.y + dy:Rumination.lag:GROUP.y + dy:Curiousity.C:GROUP.y,
                  random = ~ -1  +  dm:Curiousity.C + dy:Rumination.lag + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong2.part1,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme_mod.part2)
#lme mediation and moderation
model_lme_mod.part1 <- lme(fixed = z ~ -1 + 
                       # m as outcome
                       dm + dm:Curiousity.C + dm:Time.C + 
                       # add in group as main effect and moderator
                       dm:GROUP.y + dm:Curiousity.C:GROUP.y + 
                       
                       # y as outcome
                       dy + dy:Rumination.lag + dy:Curiousity.C + dy:Time.C +
                       #  add in group as main effect and moderator
                       dy:GROUP.y + dy:Rumination.lag:GROUP.y + dy:Curiousity.C:GROUP.y,
                  random = ~ -1  +  dm:Curiousity.C + dy:Rumination.lag + dy:Curiousity.C | ID, correlation= corAR1(),
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong2.part1,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme_mod.part1)
anova(model_lme_mod.part2, model_lme_mod.part1)

```


#### Intercepts for moderation:

* 0 = intercept in the `m` as outcome model ($d_{Mi}$: We expect this to be zero because we person-centered all the data)    

* 0 = intercept in the `y` as outcome model ($d_{Yi}$: We expect this to be zero because we person-centered all the data)

#### Effects on M (Rumination.C):

* **`r round(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C"], 3)` = effect of x --> m (*a*: work stressors predicting work dissatisfaction across averaging conditions)**     

* `r round(lme4::fixef(model_lme_mod.part1)["dm:Time.C"], 3)` = effect of time --> m (time trend in work dissatisfaction averaging across conditions)   

* `r round(lme4::fixef(model_lme_mod.part1)["dm:GROUP.y"], 3)` = effect of treatment --> m 

* **`r round(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C:GROUP.y"], 3)` = interaction of x and treatment --> m (moderation of *a*: differences in effect of work stressors predicting work dissatisfaction as a function of treatment )** 

#### Effects on Y (Stress.C):

* **`r round(lme4::fixef(model_lme_mod.part1)["dy:Rumination.C"], 3)` = effect of m --> y (*b*: work dissatisfaction predicting relationship dissatisfaction averaging across conditions)**    

* **`r round(lme4::fixef(model_lme_mod.part1)["Curiousity.C:dy"], 3)` = effect of x --> y (*c'*: work stressors predicting relationship dissatisfaction, after adjusting for work dissatisfaction, averaging across conditions)** 

* `r round(lme4::fixef(model_lme_mod.part1)["Time.C:dy"], 3)` = effect of time --> y (time trend in relationship dissatisfaction, averaging across conditions) 

* `r round(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy"], 3)` = effect of treatment --> y 

* **`r round(lme4::fixef(model_lme_mod.part1)["Curiousity.C:GROUP.y:dy"], 3)` = interaction of x and treatment --> y (moderation of *c*: differences in effect of work stressors predicting relationship dissatisfaction as a function of treatment)** 

* **`r round(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy:Rumination.C"], 3)` = interaction of m and treatment --> y (moderation of *b*: differences in effect of work dissatisfaction predicting relationship dissatisfaction as a function of treatment)** 

```{r}
#for moderation 
FE_mod <- lme4::fixef(model_lme_mod.part1)
FE_mod
```
Let's put the parameters into named objects, as these will be useful later.
```{r}

#making parameter objects

# mediation paths
a_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C"])
a_mod
b_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dy:Rumination.lag"])
b_mod
cprime_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["Curiousity.C:dy"])
cprime_mod

# moderation of mediation paths
aXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C:GROUP.y"])
aXtreat_mod
bXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy:Rumination.lag"])
bXtreat_mod
cprimeXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["Curiousity.C:GROUP.y:dy"])
cprimeXtreat_mod
```


### Random Effects
Random effects can be obtained using similar code above. Of particular interest, we will need the covariance of a and b that is left over after taking our moderator, `treat`, into account (i.e., their residual covariance).

As before, we need to convert the correlation of these random effects to a covariance.
```{r}
cortocov <- function (r, var1, var2) {
  cov=r*((var1*var2)^0.5)
  return(cov)
}
covajbj_mod <- cortocov(r = as.numeric(VarCorr(model_lme_mod.part1)["dy:Rumination.lag","Corr"]),
         var1 = as.numeric(VarCorr(model_lme_mod.part1)["dm:Curiousity.C","Variance"]),
         var2 = as.numeric(VarCorr(model_lme_mod.part1)["dy:Rumination.lag","Variance"]))
covajbj_mod
```


### Calculating mediated effect (e.g., indirect effect)  
The indirect effect is again a combination of a*b + cov(ab), but this time we have moderation on both the a and b paths to account for. 

#### Treatment Group (`treat = 1`)
First, we compute the indirect effect for the average person in the **treatment** group. 
```{r}
treat_weight <- 0.5 #reflects effect coding of treat = +.5

indirecteffect_treat_mod <- (a_mod + treat_weight*aXtreat_mod)*
                            (b_mod + treat_weight*bXtreat_mod) +
                              covajbj_mod   
indirecteffect_treat_mod
```

#### Control Group (`treat = 0`)
Next, we compute the indirect effect for the average person in the **control** group. 
```{r}
control_weight <- -0.5 #reflects effect coding of treat = -.5

indirecteffect_control_mod <- (a_mod + control_weight*aXtreat_mod)*
                              (b_mod + control_weight*bXtreat_mod) +
                              covajbj_mod   
indirecteffect_control_mod
```


#### Person-specific Mediation Estimates


```{r}
#get c 
#make x,y,m 
EMA.c <- df 
EMA.c$X <- df$Curiousity.C
EMA.c$M <- df$Stress.lag
EMA.c$Y <- df$Rumination.lag


#PART 1 restructure for just mindfulness--> rumination --> stress
MRS.data2 <- EMA.c[,c("GROUP", "ID","TIME2","Time.C","Curiousity.C","Stress.lag", "Rumination.lag","X","M","Y")]
#look at  data set
#head(MRS.data)
#tail(MRS.data)
datalong.c <- melt(data=MRS.data2,
               id.vars=c("GROUP", "ID","TIME2","Time.C","Curiousity.C","Stress.lag","Rumination.lag","X"),
               na.rm=FALSE, variable.name="dv",
               value.name="z")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#adding the double indicators 
datalong.c$dy <- ifelse(datalong.c$dv=="Y", 1, 0)
datalong.c$dm <- ifelse(datalong.c$dv=="M", 1, 0)
datalong.c$dvnum <- ifelse(datalong.c$dv=="M", 1, 0)

#reordering for convenience 
datalong.c <- datalong.c[order(datalong.c$ID,datalong.c$Time.C,datalong.c$dm),]

```


```{r}
# NEED TO RUN THE MODEL_LME.PART1 NO MODERATION FIRST !!!!!!!

#meditation FIRST
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation model
model_lme.c <- lme(fixed = z ~ -1 + 
                             dm + dm:Curiousity.C + dm:Time.C + #m as outcome
                             dy + dy:Stress.lag + dy:Curiousity.C + dy:Time.C, #y as outcome
                  random = ~ -1  +  dm:Curiousity.C + dy:Stress.lag + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.c,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.c)
```

```{r}
#starting moderation set up 
re <- data.frame(ranef(model_lme.c))
re$ID <- row.names(re)
#get group in there
group2 <- as.data.frame(datalong.c$GROUP,datalong.c$ID)
names(group2)[1] <- "GROUP"
group2$ID <- row.names(group)
group.re <- summaryBy(GROUP ~ ID, data=group2, FUN=mean, na.rm=T)
names(group.re)[2] <- "GROUP"
re$GROUP <- group.re$GROUP 
re$GROUP <- recode(re$GROUP, "0"=-0.5 , "1"=0.5 )

#don't think this line next is important? 
#re <- re %>% arrange(dm.TMS.C)

# add this new variable to the original data, resulting in a new dataframe, datalong2
datalong.2.c <- merge(datalong.c, dplyr::select(re, ID, GROUP), by = "ID")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#REDO for PART 1 restructure for mindfulness--> rumination --> well-being
#REDO for PART 1 restructure for mindfulness--> rumination --> mood
```

```{r}
#running moderation Model for Part 1 stress
#part 1
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation and moderation
model_lme.c <- lme(fixed = z ~ -1 + 
                       # m as outcome
                       dm + dm:Curiousity.C + dm:Time.C + 
                       # add in group as main effect and moderator
                       dm:GROUP.y + dm:Curiousity.C:GROUP.y + 
                       
                       # y as outcome
                       dy + dy:Stress.lag + dy:Curiousity.C + dy:Time.C +
                       #  add in group as main effect and moderator
                       dy:GROUP.y + dy:Stress.lag:GROUP.y + dy:Curiousity.C:GROUP.y,
                  random = ~ -1  +  dm:Curiousity.C + dy:Stress.lag + dy:Curiousity.C | ID,
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.2.c,
                  na.action = na.exclude)
                  #control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.c)
```