---
title: "Mediation-part1"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library (ggplot2)
library(dplyr)
library(psych)
library(plyr)
library(doBy)
library(lme4)
library(nlme)
library(jtools)
library("broom.mixed")
library(misty)
#cleaned data in excel to get day of study numbers 
#importing new data 
EMA <- readxl::read_xlsx("~/Desktop/Thesis/Aggregating-EMA/EMA-DATA-PREFINAL.xlsx")
#recode group to 0 and 1 
#0= control 1= intervention
EMA$GROUP.mean <- recode(EMA$GROUP.mean, "1"=0 , "2"=1 )
#reode day to start with 0-34
EMA$TIME <- (EMA$DAY-1)
EMA$ID <- substr(EMA$ID, 13, 15)
as.numeric(EMA$ID)
EMA$PROMIS<- round(EMA$PROMIS.mean)
EMA$STRESSOR<- round(EMA$STRESSOR.mean)
EMA$PWB <- round(EMA$PWB.mean)
EMA$Rumination <- round (EMA$Rumination.mean)
EMA$GROUP <- round(EMA$GROUP.mean)
#subset
EMA.subset <- subset.data.frame(EMA, TIME>=7)
EMA.subset$TIME2 <- (EMA.subset$TIME-7)
EMA.tx<- subset.data.frame(EMA.subset, GROUP==1)
EMA.control<- subset.data.frame(EMA.subset, GROUP==0)
```

```{r}
#clean data and import packages
library(ggplot2); theme_set(theme_bw()) #for data viz
library(lme4) #for multilevel models
library(nlme) #for multilevel models
library(psych) #for describing data
library(tidyr); library(dplyr) #for manipulating data 
library(bmlm) #for Bayesian 1-1-1 mediation
library(reshape2)
library(broom) #for converting results summary to df
# Examine first few rows of the data set
head(EMA.subset, 30)

#part 1
#x= mindfulness 
#y= stress or well-being or mood 
#m= rumination 

#part 2 
#x= mindfulness 
#y= stress or well-being or mood 
#m= sleep

#part 3
#x= mindfulness 
#y= stress or well-being or mood 
#m= exercise

#need to center variables within a person
EMA.subset$TMS.C<- center(EMA.subset$TMS.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Curiousity.C<- center(EMA.subset$Curiousity.mean, type = "CWC", group = EMA.subset$ID)
#used to check mean of ID
#describeBy(EMA.subset$TIME2, group= EMA.subset$ID==10)
EMA.subset$Decentering.C<- center(EMA.subset$Decentering.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$PROMIS.C<- center(EMA.subset$PROMIS, type = "CWC", group = EMA.subset$ID)
EMA.subset$Stress.C<- center(EMA.subset$STRESSOR, type = "CWC", group = EMA.subset$ID)
EMA.subset$PWB.C<- center(EMA.subset$PWB, type = "CWC", group = EMA.subset$ID)
EMA.subset$Rumination.C<- center(EMA.subset$Rumination, type = "CWC", group = EMA.subset$ID)
EMA.subset$Sleep.C<- center(EMA.subset$SLEEP.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$PhysicalActivity.C<- center(EMA.subset$PhysicalActivity.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Formal.C<- center(EMA.subset$FORMAL.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Informal.C<- center(EMA.subset$INFORMAL.mean, type = "CWC", group = EMA.subset$ID)
EMA.subset$Time.C<- center(EMA.subset$TIME2, type = "CWC", group = EMA.subset$ID)

#make x,y,m 
EMA.subset$X.M <- EMA.subset$Curiousity.C
EMA.subset$M.R <- EMA.subset$Rumination.C
EMA.subset$Y.W <- EMA.subset$PWB.C

#variables of interest for part 1 
vars <- c("TMS.C", "Curiousity.C", "Decentering.C", "Stress.C", "PWB.C", "PROMIS.C", "Rumination.C")
#descriptives
describe(EMA.subset[,vars])

#PART 1 restructure for just mindfulness--> rumination --> stress
MRS.data <- EMA.subset[,c("GROUP", "ID","TIME2","Time.C","Curiousity.C","PWB.C", "Rumination.C","X.M","M.R","Y.W")]
#look at  data set
#head(MRS.data)
#tail(MRS.data)
datalong.part1 <- melt(data=MRS.data,
               id.vars=c("GROUP", "ID","TIME2","Time.C","Curiousity.C","PWB.C","Rumination.C","X.M"),
               na.rm=FALSE, variable.name="dv",
               value.name="z")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#adding the double indicators 
datalong.part1$dy <- ifelse(datalong.part1$dv=="Y.W", 1, 0)
datalong.part1$dm <- ifelse(datalong.part1$dv=="M.R", 1, 0)
datalong.part1$dvnum <- ifelse(datalong.part1$dv=="M.R", 1, 0)

#reordering for convenience 
datalong.part1 <- datalong.part1[order(datalong.part1$ID,datalong.part1$Time.C,datalong.part1$dm),]

```


```{r}
# NEED TO RUN THE MODEL_LME.PART1 NO MODERATION FIRST !!!!!!!

#meditation FIRST
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation model
model_lme.part1 <- lme(fixed = z ~ -1 + 
                             dm + dm:Curiousity.C + dm:Time.C + #m as outcome
                             dy + dy:Rumination.C + dy:Curiousity.C + dy:Time.C, #y as outcome
                  random = ~ -1  +  dm:Curiousity.C + dy:Rumination.C + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.part1,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.part1)
```

Intercepts:
	0.0000 = intercept in the m as outcome model (d_Mi: We expect this to be zero because we person-centered all the data)

	0.0001 = intercept in the y as outcome model (d_Yi: We expect this to be zero (or extremely close) because we person-centered all the data)
Effects on M (Rumination):
	-0.095 = effect of x –> m (a: Mindfulness predicting Rumination) ***
	-0.007 = effect of time –> m (time trend in Rumination) ***
Effects on Y (Stress):
	-0.35 = effect of m –> y (b: Rumination predicting Well Being) ***
	0.124 = effect of x –> y (c’: Mindfulness predicting Stress, after adjusting for Rumination) ***
	0.0096 = effect of time –> y (time trend in Stress) ***

```{r}
#analyzing the mediation. 
#pulling out fixed effects info
FE.med <- fixef(model_lme.part1)
FE.med

#making parameter objects
a.med <- as.numeric(FE.med[3])
a.med
b.med <- as.numeric(FE.med[5])
b.med
cprime.med <- as.numeric(FE.med[6])
cprime.med

### Random Effects
#pulling out random effects info
VarCorr(model_lme.part1)

#The *variance* of the *a* paths
sig2_a <- as.numeric(VarCorr(model_lme.part1)["dm:Curiousity.C", "Variance"])
sig2_a

#variance of the *b* paths
sig2_b <- as.numeric(VarCorr(model_lme.part1)["dy:Rumination.C", "Variance"])
sig2_b

#variance of the *c* paths
sig2_cprime <- as.numeric(VarCorr(model_lme.part1)["Curiousity.C:dy", "Variance"])
sig2_cprime

#The residual *variance* of the mediator variable, 

sig2_em <- (1.00*as.numeric(VarCorr(model_lme.part1)["Residual", "StdDev"]))^2
sig2_em

#The residual *variance* of the outcome variable, 
sig2_ey <- (0.6968058  *as.numeric(VarCorr(model_lme.part1)["Residual", "StdDev"]))^2
sig2_ey

#The *covariance* between the $a_{j}$ and $b_{j}$ paths 
#We need to convert the correlation to a covariance, which we can by using a `cortocov` function. We can just feed in the correlation coefficient and the two variances, and it will return a covariance. 

cortocov <- function (r, var1, var2) {
  cov=r*((var1*var2)^0.5)
  return(cov)
}

covajbj <- cortocov(r = as.numeric(VarCorr(model_lme.part1)["dy:Rumination.C","Corr"]),
         var1 = as.numeric(VarCorr(model_lme.part1)["dm:Curiousity.C","Variance"]),
         var2 = as.numeric(VarCorr(model_lme.part1)["dy:Rumination.C","Variance"]))

#indirect effect (mediated effect)
indirecteffect <- a.med*b.med + covajbj   
indirecteffect

#calculating c the total effect
totaleffect <- cprime.med + a.med*b.med + covajbj 
totaleffect

#percent mediated
percentmediated <- 100*(indirecteffect/totaleffect)  
percentmediated

#percent covariance
percentcovariance <- 100*(covajbj/indirecteffect) 
percentcovariance
```

```{r}
#starting moderation set up 
re <- data.frame(ranef(model_lme.part1))
re$ID <- row.names(re)
#get group in there
group <- as.data.frame(datalong.part1$GROUP,datalong.part1$ID)
names(group)[1] <- "GROUP"
group$ID <- row.names(group)
group.re <- summaryBy(GROUP ~ ID, data=group, FUN=mean, na.rm=T)
names(group.re)[2] <- "GROUP"
re$GROUP <- group.re$GROUP 
re$GROUP <- recode(re$GROUP, "0"=-0.5 , "1"=0.5 )

#don't think this line next is important? 
#re <- re %>% arrange(dm.TMS.C)

# add this new variable to the original data, resulting in a new dataframe, datalong2
datalong2.part1 <- merge(datalong.part1, dplyr::select(re, ID, GROUP), by = "ID")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#REDO for PART 1 restructure for mindfulness--> rumination --> well-being
#REDO for PART 1 restructure for mindfulness--> rumination --> mood
```

```{r}
#running moderation Model for Part 1 stress
#part 1
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation and moderation
model_lme_mod.part1 <- lme(fixed = z ~ -1 + 
                       # m as outcome
                       dm + dm:Curiousity.C + dm:Time.C + 
                       # add in group as main effect and moderator
                       dm:GROUP.y + dm:Curiousity.C:GROUP.y + 
                       
                       # y as outcome
                       dy + dy:Rumination.C + dy:Curiousity.C + dy:Time.C +
                       #  add in group as main effect and moderator
                       dy:GROUP.y + dy:Rumination.C:GROUP.y + dy:Curiousity.C:GROUP.y,
                  random = ~ -1  +  dm:Curiousity.C + dy:Rumination.C + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong2.part1,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme_mod.part1)
```

#### Intercepts for moderation:

* 0 = intercept in the `m` as outcome model ($d_{Mi}$: We expect this to be zero because we person-centered all the data)    

* 0 = intercept in the `y` as outcome model ($d_{Yi}$: We expect this to be zero because we person-centered all the data)

#### Effects on M (Rumination.C):

* **`r round(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C"], 3)` = effect of x --> m (*a*: work stressors predicting work dissatisfaction across averaging conditions)**     

* `r round(lme4::fixef(model_lme_mod.part1)["dm:Time.C"], 3)` = effect of time --> m (time trend in work dissatisfaction averaging across conditions)   

* `r round(lme4::fixef(model_lme_mod.part1)["dm:GROUP.y"], 3)` = effect of treatment --> m 

* **`r round(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C:GROUP.y"], 3)` = interaction of x and treatment --> m (moderation of *a*: differences in effect of work stressors predicting work dissatisfaction as a function of treatment )** 

#### Effects on Y (Stress.C):

* **`r round(lme4::fixef(model_lme_mod.part1)["dy:Rumination.C"], 3)` = effect of m --> y (*b*: work dissatisfaction predicting relationship dissatisfaction averaging across conditions)**    

* **`r round(lme4::fixef(model_lme_mod.part1)["Curiousity.C:dy"], 3)` = effect of x --> y (*c'*: work stressors predicting relationship dissatisfaction, after adjusting for work dissatisfaction, averaging across conditions)** 

* `r round(lme4::fixef(model_lme_mod.part1)["Time.C:dy"], 3)` = effect of time --> y (time trend in relationship dissatisfaction, averaging across conditions) 

* `r round(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy"], 3)` = effect of treatment --> y 

* **`r round(lme4::fixef(model_lme_mod.part1)["Curiousity.C:GROUP.y:dy"], 3)` = interaction of x and treatment --> y (moderation of *c*: differences in effect of work stressors predicting relationship dissatisfaction as a function of treatment)** 

* **`r round(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy:Rumination.C"], 3)` = interaction of m and treatment --> y (moderation of *b*: differences in effect of work dissatisfaction predicting relationship dissatisfaction as a function of treatment)** 

```{r}
#for moderation 
FE_mod <- lme4::fixef(model_lme_mod.part1)
FE_mod
```
Let's put the parameters into named objects, as these will be useful later.
```{r}

#making parameter objects

# mediation paths
a_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C"])
a_mod
b_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dy:Rumination.C"])
b_mod
cprime_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["Curiousity.C:dy"])
cprime_mod

# moderation of mediation paths
aXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["dm:Curiousity.C:GROUP.y"])
aXtreat_mod
bXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["GROUP.y:dy:Rumination.C"])
bXtreat_mod
cprimeXtreat_mod <- as.numeric(lme4::fixef(model_lme_mod.part1)["Curiousity.C:GROUP.y:dy"])
cprimeXtreat_mod
```


### Random Effects
Random effects can be obtained using similar code above. Of particular interest, we will need the covariance of a and b that is left over after taking our moderator, `treat`, into account (i.e., their residual covariance).

As before, we need to convert the correlation of these random effects to a covariance.
```{r}
cortocov <- function (r, var1, var2) {
  cov=r*((var1*var2)^0.5)
  return(cov)
}
covajbj_mod <- cortocov(r = as.numeric(VarCorr(model_lme_mod.part1)["dy:Rumination.C","Corr"]),
         var1 = as.numeric(VarCorr(model_lme_mod.part1)["dm:Curiousity.C","Variance"]),
         var2 = as.numeric(VarCorr(model_lme_mod.part1)["dy:Rumination.C","Variance"]))
covajbj_mod
```


### Calculating mediated effect (e.g., indirect effect)  
The indirect effect is again a combination of a*b + cov(ab), but this time we have moderation on both the a and b paths to account for. 

#### Treatment Group (`treat = 1`)
First, we compute the indirect effect for the average person in the **treatment** group. 
```{r}
treat_weight <- 0.5 #reflects effect coding of treat = +.5

indirecteffect_treat_mod <- (a_mod + treat_weight*aXtreat_mod)*
                            (b_mod + treat_weight*bXtreat_mod) +
                              covajbj_mod   
indirecteffect_treat_mod
```

#### Control Group (`treat = 0`)
Next, we compute the indirect effect for the average person in the **control** group. 
```{r}
control_weight <- -0.5 #reflects effect coding of treat = -.5

indirecteffect_control_mod <- (a_mod + control_weight*aXtreat_mod)*
                              (b_mod + control_weight*bXtreat_mod) +
                              covajbj_mod   
indirecteffect_control_mod
```


#### Person-specific Mediation Estimates

```{r}

ind_ranefs <- data.frame(coef(model_lme_mod.part1))
ind_ranefs$ID <- row.names(ind_ranefs)

# combine with info about treatment condition
ind_ranefs_treat <- merge(ind_ranefs, dplyr::select(re, ID, GROUP), by = "ID")


# compute person specific indirect effect
ind_ranefs_treat$indirect <-
  # moderated a path
  (ind_ranefs_treat$dm.Curiousity.C + ind_ranefs_treat$GROUP * ind_ranefs_treat$dm.Curiousity.C.GROUP.y) *
  # moderated b path
  (ind_ranefs_treat$dy.Rumination.C + ind_ranefs_treat$GROUP * ind_ranefs_treat$GROUP.y.dy.Rumination.C) 

```


### Visualizing Moderated 1-1-1 Mediation Effect

```{r}
# create an empty y-variable strictly for plotting purposes
ind_ranefs_treat$y <- NA


ggplot(ind_ranefs_treat, aes(x = indirect, y = y, color = as.factor(GROUP))) +
  # add in data points showing subject-specific mediation effect, colored by treat
  geom_point(position = position_jitter(h = .03), alpha = .4, size = 2) + 
  
  # add in large points to indicate the mediation effect for 
  # the average person in each group
  geom_point(aes(x = indirecteffect_control_mod, y = y), color = "darkred", size = 6) + 
  geom_point(aes(x = indirecteffect_treat_mod, y = y), color = "navyblue", size = 6) +
  
  scale_color_manual(values = c("red", "blue"), labels = c("Control", "Treatment")) +
  theme_bw()+
  xlab("Indirect Effect") +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank()) +
  labs(color = "Intervention Condition") +
  theme(legend.position="bottom")
``` 

As this figure shows, the 1-1-1 mediation effect differs between conditions. The average person in the treatment condition showed an attenuated indirect effect of daily work stress on relationship dissatisfaction via work dissatisfaction, compared to those in the control condition. Note that we still see some unexplained between-person heterogeneity, as indicated by the smaller dots in the figure above. 


For further information and discussion of moderated mediation in multilevel models, you may wish to check out this paper by Bauer, Preacher, and Gil: 
[http://quantpsy.org/pubs/bauer_preacher_gil_2006.pdf](http://quantpsy.org/pubs/bauer_preacher_gil_2006.pdf). There are some interesting extensions that can be used to obtain bootstrapped confidence intervals for the mediated effects. 

```{r}
#get c 
#make x,y,m 
EMA.c <- EMA.subset
EMA.c$X <- EMA.c$Curiousity.C
EMA.c$M <- EMA.c$PWB.C
EMA.c$Y <- EMA.c$Rumination.C

#variables of interest for part 1 
vars <- c("TMS.C", "Decentering.C", "Curiousity.C", "Stress.C", "PWB.C", "PROMIS.C", "Rumination.C")
#descriptives
describe(EMA.c[,vars])

#PART 1 restructure for just mindfulness--> rumination --> stress
MRS.data <- EMA.c[,c("GROUP", "ID","TIME2","Time.C","Curiousity.C","PWB.C", "Rumination.C","X","M","Y")]
#look at  data set
#head(MRS.data)
#tail(MRS.data)
datalong.c <- melt(data=MRS.data,
               id.vars=c("GROUP", "ID","TIME2","Time.C","Curiousity.C","PWB.C","Rumination.C","X"),
               na.rm=FALSE, variable.name="dv",
               value.name="z")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#adding the double indicators 
datalong.c$dy <- ifelse(datalong.c$dv=="Y", 1, 0)
datalong.c$dm <- ifelse(datalong.c$dv=="M", 1, 0)
datalong.c$dvnum <- ifelse(datalong.c$dv=="M", 1, 0)

#reordering for convenience 
datalong.c <- datalong.c[order(datalong.c$ID,datalong.c$Time.C,datalong.c$dm),]

```


```{r}
# NEED TO RUN THE MODEL_LME.PART1 NO MODERATION FIRST !!!!!!!

#meditation FIRST
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation model
model_lme.c <- lme(fixed = z ~ -1 + 
                             dm + dm:Curiousity.C + dm:Time.C + #m as outcome
                             dy + dy:PWB.C + dy:Curiousity.C + dy:Time.C, #y as outcome
                  random = ~ -1  +  dm:Curiousity.C + dy:PWB.C + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.c,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.c)
```

```{r}
#starting moderation set up 
re <- data.frame(ranef(model_lme.c))
re$ID <- row.names(re)
#get group in there
group2 <- as.data.frame(datalong.c$GROUP,datalong.c$ID)
names(group2)[1] <- "GROUP"
group2$ID <- row.names(group)
group.re <- summaryBy(GROUP ~ ID, data=group2, FUN=mean, na.rm=T)
names(group.re)[2] <- "GROUP"
re$GROUP <- group.re$GROUP 
re$GROUP <- recode(re$GROUP, "0"=-0.5 , "1"=0.5 )

#don't think this line next is important? 
#re <- re %>% arrange(dm.TMS.C)

# add this new variable to the original data, resulting in a new dataframe, datalong2
datalong.2.c <- merge(datalong.c, dplyr::select(re, ID, GROUP), by = "ID")
#look at updated data set
#head(datalong.part1)
#tail(datalong.part1)

#REDO for PART 1 restructure for mindfulness--> rumination --> well-being
#REDO for PART 1 restructure for mindfulness--> rumination --> mood
```

```{r}
#running moderation Model for Part 1 stress
#part 1
#x= mindfulness (fwkstrcw in Bolger)
#y= stress (frediscw in Bolger)
#m= rumination (fwkdiscw in Bolger)

#lme mediation and moderation
model_lme.c <- lme(fixed = z ~ -1 + 
                       # m as outcome
                       dm + dm:Curiousity.C + dm:Time.C + 
                       # add in group as main effect and moderator
                       dm:GROUP.y + dm:Curiousity.C:GROUP.y + 
                       
                       # y as outcome
                       dy + dy:PWB.C + dy:Curiousity.C + dy:Time.C +
                       #  add in group as main effect and moderator
                       dy:GROUP.y + dy:PWB.C:GROUP.y + dy:Curiousity.C:GROUP.y,
                  random = ~ -1  +  dm:Curiousity.C + dy:PWB.C + dy:Curiousity.C | ID, 
                  weights = varIdent(form = ~ 1 | dvnum), #separate sigma^{2}_{e} for each outcome
                  data = datalong.2.c,
                  na.action = na.exclude,
                  control = lmeControl(opt = "optim", maxIter = 200, msMaxIter = 200, niterEM = 50, msMaxEval = 400))

summary(model_lme.c)
```